Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib

### 2023-09-20 texor.Rmd
handled two of the issues

In 2019, with the help of funding from the R Consortium, work began to update operations. One aspect of this was to change from LaTeX article submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf. There are numerous benefits of HTML format:

@@

The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility. Thus it only creates the R Markdown and an HTML version leaving the original PDF as published. This will also be used to produce HTML versions of newly accepted papers only submitted as LaTeX.

However, it could be useful to help authors convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.

One of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element. In an R Markdown code chunk generating a figure, this can be set using `fig.alt`. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

>>>>

In 2019, with the help of funding from the R Consortium, work began to update operations to adopt an R Markdown template, that would produce both HTML and PDF versions of the paper. In mind, was also to develop, adopt and encourage good standards for R Journal articles.

There are numerous benefits of HTML format:

@@

The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility. Thus it only creates an `.Rmd` (R Markdown) file which is used to produce only an HTML version. The original PDF remains as published. The goals is use use the packages for future articles where the authors have chosen to submit only in the legacy LaTeX template.

However, the software can also be used by authors themselves to convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.

Adopting the new R Markdown template also provides better standards for R Journal articles. It encourages authors to be more conscious that their work is reproducible. There is more standardization of file names and file structure, and this helps the editors and reviewers check papers. This in turn makes it easier to build the issues eliminating some common errors that authors make that cause compilation failure.

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element. In an R Markdown code chunk generating a figure, this can be set using `fig.alt`. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

There are many other journals that have a similar backlog of legacy articles. A current and ongoing project is being conducted by @arxiv to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML [@latexML]. The work conducted for the R Journal might of interest to similar small journals that operate without paid staff. Converting legacy articles provides accessibility to a wealth of educational material.

### 2023-09-20 texor.tex
Bumped the rebib version with the CRAN version.
handled two of the issues

The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility. Thus it only creates the R Markdown and an HTML version leaving the original PDF as published. This will also be used to produce HTML versions of newly accepted papers only submitted as LaTeX.

However, it could be useful to help authors convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.

One of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element. In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

@@

\begin{itemize}
\tightlist
\item
  \CRANpkg{texor}: Version 1.2.0 \url{https://abhi-1u.github.io/texor}
\item
  \CRANpkg{rebib}: Version 0.3.0 \url{https://abhi-1u.github.io/rebib/}
\item
  This paper: \url{https://github.com/Abhi-1U/texor-rjarticle}
\item
  More details on \CRANpkg{rjtools} are at \url{https://rjournal.github.io/rjtools/}
\end{itemize}

@@

@@

>>>>

The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility. Thus it only creates an \texttt{.Rmd} (R Markdown) file which is used to produce only an HTML version. The original PDF remains as published. The goals is use use the packages for future articles where the authors have chosen to submit only in the legacy LaTeX template.

However, the software can also be used by authors themselves to convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.

Adopting the new R Markdown template also provides better standards for R Journal articles. It encourages authors to be more conscious that their work is reproducible. There is more standardization of file names and file structure, and this helps the editors and reviewers check papers. This in turn makes it easier to build the issues eliminating some common errors that authors make that cause compilation failure.

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element. In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

There are many other journals that have a similar backlog of legacy articles. A current and ongoing project is being conducted by {``HTML as an Accessible Format for Papers''} (2023) to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML ({``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter''} 2023). The work conducted for the R Journal might of interest to similar small journals that operate without paid staff. Converting legacy articles provides accessibility to a wealth of educational material.

@@

\begin{itemize}
\tightlist
\item
  \CRANpkg{texor}: Version 1.2.0 \url{https://abhi-1u.github.io/texor}
\item
  \CRANpkg{rebib}: Version 0.3.2 \url{https://abhi-1u.github.io/rebib/}
\item
  This paper: \url{https://github.com/Abhi-1U/texor-rjarticle}
\item
  More details on \CRANpkg{rjtools} are at \url{https://rjournal.github.io/rjtools/}
\end{itemize}

@@

\leavevmode\vadjust pre{\hypertarget{ref-arxiv}{}}%
{``HTML as an Accessible Format for Papers.''} 2023. https://info.arxiv.org/about/accessible\_HTML.html.

@@

\leavevmode\vadjust pre{\hypertarget{ref-latexML}{}}%
{``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter.''} 2023. https://math.nist.gov/~BMiller/LaTeXML/.
